Allele Frequencies Common Gardens

### 2020-02-06 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
Analyze genomic distribution of winter survival selected SNPs

18,333,684

PKLE KBSM CLMB >= 25
```{r}
big_three <- snps %>%
  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25)
saveRDS(big_three, "SNPs_above_MAC_25_in_big_three_sites.rds")
```

@@

12,358,701 SNPs

PKLE TMPL BRKG KBSM LINC CLMB STIL >= 25
```{r}
fourway_seven <- snps %>%
  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
           MAC_LINC_2019 >= 25, MAC_LINC_2018 >= 25,
           MAC_STIL_2019 >= 25, MAC_STIL_2018 >= 25,
           MAC_TMPL_2019 >= 25, MAC_TMPL_2018 >= 25,
           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
           MAC_BRKG_2019 >= 25, MAC_BRKG_2018 >= 25,
           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25)
saveRDS(fourway_seven, file = "SNPs_above_MAC_25_in_fourway_7_sites.rds")
```

@@

1,963,771 SNPs
```{r}
missing_frz <- snps %>%
  filter(MAC_FRMI_2018 >= 25 & MAC_LINC_2018 >= 25 & MAC_CLMB_2018 >= 25 &
           MAC_BRKG_2018 >= 25, MAC_FRMI_2019 < 15 | MAC_LINC_2018 < 25 | 
           MAC_CLMB_2018 < 25 | MAC_BRKG_2019 < 25) %>%
  mutate(avediff = ((MAC_FRMI_2018 - MAC_FRMI_2019) + (MAC_LINC_2018 - MAC_LINC_2019) + (MAC_CLMB_2018 - MAC_CLMB_2019) + (MAC_BRKG_2018 - MAC_BRKG_2019))/4)

saveRDS(missing_frz, file = "Missing_SNPs_2019_4_freezing_sites.rds")
```

@@

```{r}
missing_frz %>%
  ggplot(aes(x = POS, y = avediff)) + 
  geom_point() + 
  facet_wrap(~CHR)
save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites.png", plot = last_plot())

missing_frz %>%
  ggplot(aes(x = POS, y = avediff)) + 
  geom_hex() + 
  facet_wrap(~CHR) + 
  scale_color_viridis() + 
  scale_fill_viridis()
save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites_hex.png", plot = last_plot())
```

>>>>

18,704,908 SNPs

PKLE KBSM CLMB >= 25
```{r}
big_three <- snps %>%
  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25)
saveRDS(big_three, "SNPs_above_MAC_25_in_big_three_sites.rds")
10183856+6027776+2493276
```

@@

12,677,132 SNPs

PKLE TMPL BRKG KBSM LINC CLMB STIL >= 25
```{r}
fourway_seven <- snps %>%
  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
           MAC_LINC_2019 >= 25, MAC_LINC_2018 >= 25,
           MAC_STIL_2019 >= 25, MAC_STIL_2018 >= 25,
           MAC_TMPL_2019 >= 25, MAC_TMPL_2018 >= 25,
           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
           MAC_BRKG_2019 >= 25, MAC_BRKG_2018 >= 25,
           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25)
saveRDS(fourway_seven, file = "SNPs_above_MAC_25_in_fourway_7_sites.rds")
10183856+2493276
```

## microbiome

Just got the results from PKLE thus far.

21,525,180 SNPs

```{r}
microbiome <- snps %>%
  filter(MAC_PKLE_2019 >= 25)
10183856+6027776+2493276+2820272
```

@@

2,288,528 SNPs
```{r}
missing_frz <- snps %>%
  filter(MAC_FRMI_2018 >= 25 & MAC_LINC_2018 >= 25 & MAC_CLMB_2018 >= 25 &
           MAC_BRKG_2018 >= 25, MAC_FRMI_2019 < 15 | MAC_LINC_2019 < 25 | 
           MAC_CLMB_2019 < 25 | MAC_BRKG_2019 < 25) %>%
  mutate(avediff = ((MAC_FRMI_2018 - MAC_FRMI_2019) + (MAC_LINC_2018 - MAC_LINC_2019) + (MAC_CLMB_2018 - MAC_CLMB_2019) + (MAC_BRKG_2018 - MAC_BRKG_2019))/4)

saveRDS(missing_frz, file = "Missing_SNPs_2019_4_freezing_sites.rds")

25/(615*2)
25/(496*2)
```

```{r}
snp_subsets <- snps %>%
  mutate(SNP_tensite_twoyear = ifelse(MAC_PKLE_2019 >= 25 & 
                                        MAC_PKLE_2018 >= 25 & 
                                        MAC_FRMI_2019 >= 15 & 
                                        MAC_FRMI_2018 >= 25 & 
                                        MAC_OVTN_2019 >= 25 & 
                                        MAC_OVTN_2018 >= 25 & 
                                        MAC_LINC_2019 >= 25 & 
                                        MAC_LINC_2018 >= 25 & 
                                        MAC_STIL_2019 >= 25 & 
                                        MAC_STIL_2018 >= 25 & 
                                        MAC_TMPL_2019 >= 25 & 
                                        MAC_TMPL_2018 >= 25 & 
                                        MAC_CLMB_2019 >= 25 & 
                                        MAC_CLMB_2018 >= 25 & 
                                        MAC_BRKG_2019 >= 25 & 
                                        MAC_BRKG_2018 >= 25 & 
                                        MAC_KBSM_2019 >= 25 & 
                                        MAC_KBSM_2018 >= 25 & 
                                        MAC_KING_2019 >= 25 & 
                                        MAC_KING_2018 >= 25, "Y", "N"),
         SNP_big_three = ifelse(MAC_PKLE_2019 >= 25 & 
                                        MAC_PKLE_2018 >= 25 & 
                                        MAC_CLMB_2019 >= 25 & 
                                        MAC_CLMB_2018 >= 25 & 
                                        MAC_KBSM_2019 >= 25 & 
                                        MAC_KBSM_2018 >= 25, "Y", "N"),
         SNP_fourway_seven = ifelse(MAC_PKLE_2019 >= 25 & 
                                        MAC_PKLE_2018 >= 25 & 
                                        MAC_LINC_2019 >= 25 & 
                                        MAC_LINC_2018 >= 25 & 
                                        MAC_STIL_2019 >= 25 & 
                                        MAC_STIL_2018 >= 25 & 
                                        MAC_TMPL_2019 >= 25 & 
                                        MAC_TMPL_2018 >= 25 & 
                                        MAC_CLMB_2019 >= 25 & 
                                        MAC_CLMB_2018 >= 25 & 
                                        MAC_BRKG_2019 >= 25 & 
                                        MAC_BRKG_2018 >= 25 & 
                                        MAC_KBSM_2019 >= 25 & 
                                        MAC_KBSM_2018 >= 25, "Y", "N"),
         SNP_microbiome = ifelse(MAC_PKLE_2019 >= 25 & 
                                        MAC_PKLE_2018 >= 25, "Y", "N"),
         SNP_missing_frz = ifelse((MAC_FRMI_2018 >= 25 & MAC_BRKG_2018 >= 25 &
                                     MAC_LINC_2018 >= 25 & 
                                     MAC_CLMB_2018 >= 25) &
                                     (MAC_LINC_2019 < 25 | 
                                        MAC_CLMB_2019 < 25 | 
                                        MAC_BRKG_2019 < 25 |
                                    MAC_FRMI_2019 < 15), "Y", "N")
         )

snp_subsets %>%
  dplyr::select(CHR, POS, starts_with("SNP_")) %>%
  saveRDS(., file = "SNP_subsets_33M_markers.rds")

missing_frz <- snp_subsets %>%
  filter(SNP_missing_frz == "Y")

snp_subsets %>%
  filter(SNP_missing_frz == "Y") %>%
  saveRDS(., file = "Missing_SNPs_2019_4_freezing_sites.rds")

snp_subsets %>%
  group_by(SNP_missing_frz) %>% tally()
snp_subsets %>%
  group_by(SNP_tensite_twoyear, SNP_big_three, SNP_fourway_seven, SNP_microbiome) %>% tally()
```

@@

```{r}
theme_set(theme_poster)
missing_frz %>%
  ggplot(aes(x = POS, y = avediff)) + 
  geom_point() + 
  facet_wrap(~CHR)
save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites.png", plot = last_plot())

missing_frz %>%
  ggplot(aes(x = POS, y = avediff)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss, 4 sites") +
  theme(legend.position = "right")
save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_FRMI_2018 >= 25 & MAC_LINC_2018 >= 25 & MAC_CLMB_2018 >= 25 &
           MAC_BRKG_2018 >= 25 & MAC_FRMI_2019 < 15 & MAC_LINC_2019 < 25 & 
           MAC_CLMB_2019 < 25 & MAC_BRKG_2019 < 25) %>%
  ggplot(aes(x = POS, y = avediff)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss, 4 sites") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_all_four_freezing_sites_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_BRKG_2018 >= 25 & MAC_BRKG_2019 < 15) %>%
  mutate(Allele_loss = MAC_BRKG_2018 - MAC_BRKG_2019) %>%
  ggplot(aes(x = POS, y = Allele_loss)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_BRKG_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_LINC_2018 >= 25 & MAC_LINC_2019 < 15) %>%
  mutate(Allele_loss = MAC_LINC_2018 - MAC_LINC_2019) %>%
  ggplot(aes(x = POS, y = Allele_loss)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_LINC_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_FRMI_2018 >= 25 & MAC_FRMI_2019 < 15) %>%
  mutate(Allele_loss = MAC_FRMI_2018 - MAC_FRMI_2019) %>%
  ggplot(aes(x = POS, y = Allele_loss)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_FRMI_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_CLMB_2018 >= 25 & MAC_CLMB_2019 < 15) %>%
  mutate(Allele_loss = MAC_CLMB_2018 - MAC_CLMB_2019) %>%
  ggplot(aes(x = POS, y = Allele_loss)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_CLMB_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz <- missing_frz %>%
  mutate(Loss_LINC = MAC_LINC_2018 - MAC_LINC_2019,
         Loss_BRKG = MAC_BRKG_2018 - MAC_BRKG_2019,
         Loss_FRMI = MAC_FRMI_2018 - MAC_FRMI_2019,
         Loss_CLMB = MAC_CLMB_2018 - MAC_CLMB_2019)

missing_frz %>%
  group_by(Loss_BRKG > 75, Loss_LINC > 78, Loss_FRMI > 54, Loss_CLMB > 51) %>%
  tally() %>%
  write_csv(., path = "SNPs_loss_overlaps_1percent_tail_2019_four_freezing_sites.csv")

plot1 <- missing_frz %>%
  ggplot(aes(x = Loss_CLMB)) + geom_histogram(binwidth = 2)
save_plot(filename = "Histogram_SNPs_lost_CLMB.png", plot = plot1)
plot1 <- missing_frz %>%
  ggplot(aes(x = Loss_FRMI)) + geom_histogram(binwidth = 2)
save_plot(filename = "Histogram_SNPs_lost_FRMI.png", plot = plot1)

plot1 <- missing_frz %>%
  ggplot(aes(x = Loss_BRKG)) + geom_histogram(binwidth = 2)
save_plot(filename = "Histogram_SNPs_lost_BRKG.png", plot = plot1)

plot1 <- missing_frz %>%
  ggplot(aes(x = Loss_LINC)) + geom_histogram(binwidth = 2)
save_plot(filename = "Histogram_SNPs_lost_LINC.png", plot = plot1)

missing_frz %>%
  summarise(top_LINC = quantile(Loss_LINC, 0.99),
            top_BRKG = quantile(Loss_BRKG, 0.99),
            top_FRMI = quantile(Loss_FRMI, 0.99),
            top_CLMB = quantile(Loss_CLMB, 0.99),
            )

# LINC & BRKG are the most similar

all_frz <- snp_subsets %>%
  mutate(Loss_LINC = (MAC_LINC_2018 - MAC_LINC_2019)/MAC_LINC_2018,
         Loss_BRKG = (MAC_BRKG_2018 - MAC_BRKG_2019)/MAC_BRKG_2018,
         Loss_FRMI = (MAC_FRMI_2018 - MAC_FRMI_2019)/MAC_FRMI_2018,
         Loss_CLMB = (MAC_CLMB_2018 - MAC_CLMB_2019)/MAC_CLMB_2018)
all_frz %>%
  summarise(top_LINC = quantile(Loss_LINC, 0.99, na.rm = TRUE),
            top_BRKG = quantile(Loss_BRKG, 0.99, na.rm = TRUE),
            top_FRMI = quantile(Loss_FRMI, 0.99, na.rm = TRUE),
            top_CLMB = quantile(Loss_CLMB, 0.99, na.rm = TRUE)
            )
all_frz %>%
  group_by(Loss_BRKG > 159, Loss_LINC > 179, Loss_FRMI > 84, Loss_CLMB > 94) %>%
  tally() %>%
  write_csv(., path = "All_SNPs_lost_alleles_1percent_tail_2019_four_freezing_sites.csv")

missing_frz %>%
  filter(Loss_BRKG > 75 | Loss_LINC > 78 | Loss_FRMI > 54 | Loss_CLMB > 51) %>% saveRDS(., file = "Lost_SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
all_frz %>%
  filter(Loss_BRKG > 159 | Loss_LINC > 179 | Loss_FRMI > 84 | Loss_CLMB > 94) %>% saveRDS(., file = "SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
```

  top_LINC top_BRKG top_FRMI top_CLMB
     <dbl>    <dbl>    <dbl>    <dbl>
1       78       75       54       51

# Frequencies lost 
```{r}
frz_freq <- readRDS("SNPs_with_1percent_tail_frequency_change_freezing_sites.rds") 
loss_freq <- readRDS("Lost_SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
```

```{r}
loss_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_LINC/MAC_LINC_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_LINC.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_LINC/MAC_LINC_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_LINC.png", plot = last_plot(), base_asp = 4, base_height = 4)
```

```{r}
loss_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_CLMB/MAC_CLMB_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_CLMB.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_CLMB/MAC_CLMB_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_CLMB.png", plot = last_plot(), base_asp = 4, base_height = 4)

loss_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_FRMI/MAC_FRMI_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_FRMI.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_FRMI/MAC_FRMI_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_FRMI.png", plot = last_plot(), base_asp = 4, base_height = 4)

loss_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_BRKG/MAC_BRKG_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_BRKG.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_BRKG/MAC_BRKG_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_BRKG.png", plot = last_plot(), base_asp = 4, base_height = 4)
```

```{r}
loss_freq %>%
    filter(Loss_BRKG > 75 | Loss_LINC > 78 | Loss_FRMI > 54 | 
             Loss_CLMB > 51) %>% 
  mutate(Loss_freq = (Loss_BRKG/MAC_BRKG_2018 + Loss_FRMI/MAC_FRMI_2018 +
                            Loss_LINC/MAC_LINC_2018 +
                            Loss_CLMB/MAC_CLMB_2018)/4) %>% 
  ggplot(aes(x = POS/1000000, y = Loss_freq)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_one_lost_all_four_sites.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
  filter(Loss_BRKG > 159 | Loss_LINC > 179 | Loss_FRMI > 84 | 
           Loss_CLMB > 94) %>%
  mutate(Loss_freq = (Loss_BRKG/MAC_BRKG_2018 + Loss_FRMI/MAC_FRMI_2018 +
                            Loss_LINC/MAC_LINC_2018 +
                            Loss_CLMB/MAC_CLMB_2018)/4) %>% 
  ggplot(aes(x = POS/1000000, y = Loss_freq)) +
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_all_four_sites.png", plot = last_plot(), base_asp = 4, base_height = 4)
```
